Universals of word order result from optimization of grammars for efficient communication

### 2019-02-24 second-draft/4-paper-names.tex
visualization

\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}

@@

%\section{DLM}

%
%\begin{figure}
%    \centering
%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure. In this example, Grammar 1 achieves the lowest overall dependency length. Indeed, this is the }
%	\label{fig:pareto-plane}
%\end{figure}
%
%

@@

@@

We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
%This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
%Similarly, most of the other dependencies in Dryer's formulation can be formalized:
%for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
%Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).

Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
Further collapsing correlations that cannot be formalized individually in UD, we otained 10 formalized correlations from Dryer's 15 correlations.
%Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD. From Dryer's 15 correlations, we obtained 10 formalized correlations, roughly covering nine of Greenberg's original universals. %TODO update to 10? What happened to the 11th? Was that amod? [The 11th was adjective+standard of comparison, which is merged into obl] Maybe we could include amod as a "true rejection" of efficiency, but a false alarm for DepL

@@

@@

@@

Dependency Length:

>>>>

@@

\section{DLM}

\begin{figure*}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/four-grammars-dlm.png}
	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure. In this example, Grammar 1 achieves the lowest overall dependency length. Indeed, this is the }
	\label{fig:dlm-grammars}
\end{figure*}

@@

Dependency Length Minimization is heuristic and stipulative.

@@

\subsection{Greenbergian Word Order Correlations}
\label{sec:greenberg}
\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP       
	&   \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf}   
	&    \textbf{81}$^{***}$    
	&   \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   
	&    \textbf{76}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}  \\
copula    &    NP       
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf}
	&    \textbf{81}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}
	&    \textbf{79}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}  \\
auxiliary    &    VP       
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf} 
	&    \textbf{74}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf} 
	&    55    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  \\
noun    &    genitive      
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf} 
	&    \textbf{82}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf} 
	&    \textbf{74}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  \\
noun    &    relative clause       
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf} 
	&    \textbf{85}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf} 
	&    \textbf{77}$^{**}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  \\
complementizer    &    S        
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf} 
	&    \textbf{85}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf} 
	&    \textbf{80}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  \\
verb    &    PP     
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf} 
	&    \textbf{78}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf} 
	&    59    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  \\
want    &    VP      
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf} 
	&    \textbf{90}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf} 
	&    \textbf{92}$^{***}$    
	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf} 
  \\
 \hline
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
%	For clarity, we organize the correlations by the category of the `verb patterner'.
	For each correlation, we give our operationalization in terms of UD. %A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation. 
	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
\end{table*}

We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.

Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations from Dryer's 15 correlations.

@@

\paragraph{Results}
Results are shown in Table~\ref{table:results-dryer}.
All eight correlations are confirmed in the models estimated from the real orderings. 
We will discuss these exceptions further below.

Confirming previous theoretical arguments, we see that optimizing for dependency length correctly accounts for all word order correlations. 
Predictability and parseability predict four and six correlations, respectively, making complementary predictions.
Efficiency correctly predicts all the word order correlations.

@@

In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars. Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces the Greenbergian correlations to the extent that they can be tested on the basis of available data.

@@
